Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/algorithms.tex
on inference part 1

\begin{equation*}
\texttt{a} \mapsto \{[0,0]\times[0,0], \;\; [-1,0]\times[0,0], \;\;
[0,0]\times[-1,0], \;\; [0,0]\times[1,0], \;\; [0,1]\times[0,0]\}
\end{equation*}
%
We outline here the procedures for checking conformance
of source code against specifications (\Cref{subsec:checking})
and for inferring specifications from code (\Cref{subsec:inference}).
Both rely on a program analysis that converts array subscripts
 into sets of index schemes. We outline this analysis
first (\cref{subsec:analysis}). Note that the analysis
can be made arbitrarily more complex and wide-ranging independent
of the checking and inference procedures. At the moment, the analysis
is largely \emph{syntactic}, with only a small amount of
semantic interpretation of the code.

@@

\subsection{Infering specification automatically}
\label{subsec:inference}

We provide an inference procedure for generating specifications from
code, which are inserted automatically as comments. This helps support
maintenance of legacy code, and aids understanding of our
specification language. We illustrate inference using the
five-point stencil of \Cref{exm:checking}.

The preceding source code analysis converts the concrete syntax of array
subscripts into a subscript space in $\mathcal{P}(\bz)^N$.
Let $U$ range over finite maps from array variables to sets of
of subscript schemes.

 The
inference then attempts to construct a $\emph{Region}_N$ from this.

 For
each array variable $a \in \mathsf{dom}(U)$, the algorithm constructs a set of
$n$-dimensional rectangles covering all contiguous groups of schemes
in $U(a)$.

\begin{enumerate}[leftmargin=1.5em]
\item
A covering of (possibly overlapping) regions is computed iteratively
by the following steps until a fixed-point is reached:

\begin{enumerate}
\item \textbf{Coalesce intervals}
Intervals which are adjacent (contiguous in one dimension) are
coalesced.

\begin{defn}{[Contiguous]}
  An index scheme
  A subscript $S$ is said to be \emph{adjacent at index $j$} of another
  subscript $S'$ if for every index $i$ except $j$, $\pi_i(S) = \pi_i(S')$ but
  at $j$ then $\pi_j(S) \neq \pi_j(S')$ such that $|S| = |S'| + 1$.
\end{defn}
%
\item \textbf{Remove completely overlapped intervals}
\item \textbf{Covering}
\end{enumerate}
\end{enumerate}

>>>>

%
We outline here the procedures for checking conformance
of source code against specifications (\Cref{subsec:checking})
and for inferring specifications from code (\Cref{subsec:inference}).
Both rely on a program analysis that converts array subscripts
 into sets of index schemes. We outline this analysis
first (\cref{subsec:analysis}). Note that the analysis
can be made arbitrarily more complex and wide-ranging independent
of the checking and inference procedures. At the moment, the analysis
is largely \emph{syntactic}, with only a small amount of
semantic interpretation of the code.

@@

\subsection{Infering specification automatically}
\label{subsec:inference}
%
We provide an inference procedure for generating specifications from
code, which are inserted automatically as comments. This helps support
maintenance of legacy code, and aids understanding of our
specification language. We illustrate inference using the
five-point stencil of \Cref{exm:checking}.

The preceding source code analysis converts the concrete syntax of
array subscripts into a set of subscript schemes in
$\mathcal{P}(\bz)^N$. These subscripts are then coalesced interatively
into sets of index schemes. We denote sets of index schemes by
$\mathcal{S}$, and initial set of subscript schemes from the analysis
as $\mathcal{S}_0$. For our example:
%
\begin{equation*}
\mathcal{S}_0 = \{\{0\} \times \{0\}, \{-1\} \times \{0\},
\{0\} \times \{-1\}, \{0\} \times \{1\}, \{1\} \times \{0\}\}
\end{equation*}

%For each array variable $a \in \mathsf{dom}(U)$, the algorithm
%constructs a set of $n$-dimensional rectangles covering all contiguous
%groups of schemes in $U(a)$.

\begin{enumerate}[leftmargin=1.5em]
\item (\textbf{Covering})
A covering of (possibly overlapping) the subscript space
is calculated as a value of $Region_N$. This is computed iteratively
by the following steps until a fixed-point is reached:

\begin{enumerate}[leftmargin=0.5em]
\item (\textbf{Coalesce intervals})
%
Index schemes which are \emph{contiguous} in one dimension are
coalesced.

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}{[Contiguous]}
Two index schemes $S$ and $T$ are \emph{contiguous}
written $\contig{S}{T}$,
iff $\pi_i(S) = \pi_i(T)$
for all $1 \leq i \leq N$ apart from some dimension $k$ such that
$\pi_k(S) = [a, b]$ and $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
\begin{align*}
\mathsf{coalesce}(\mathcal{S}) = \; &
\{\,S \cup T \mid
\forall S \in \mathcal{S}, \;
\forall T \in \mathcal{S}. \;
\contig{S}{T} \} \\
\cup \; &
\{ \,\;\;\; S \;\;\; \mid \forall S \in \mathcal{S},
\; \not \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
%
That is, the new set of index schemes $\mathsf{coalesc}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.
\item (\textbf{Remove redundant schemes})
Any indexing schemes which are subsumed by another are filtered out:
%
\begin{align*}
\textsf{nonOverlap}(\mathcal{S}) =
\{S \mid \forall S \in \mathcal{S}, \; \not \exists T \in (\mathcal{S} -
  \{S\}), \; \forall i . \; \pi_i(S) \subseteq \pi_i(T) \}
\end{align*}
%
\dnote{I'm starting to suspect that this step is actually redundant
  now- but I will need to do some math to check}
%\item \textbf{Covering}
\end{enumerate}
The least fixed-point of $\textsf{nonOverlap} \circ \textsf{coalesce}$
is computed, providing a minimal indexing scheme covering the space.
\dnote{Proof?}

For our example:
%
\begin{align*}
\textsf{coalesce}(\mathcal{S}_0) & =
\{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times [0,1],\,[0,1]
  \times [0,0]\} \\
%\textsf{nonOverlap}(\textsf{coalesce}(\mathcal{S}_0)) & =
%                                                        \textsf{coalesce}(\mathcal{S}_0)
%  \\
\textsf{coalesce}^2(\mathcal{S}_0) & =
\{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\}
\end{align*}
%
The fixed-pointed is reached within two applications of
\textsf{coalesce} for our example.
\dnote{We probably need to say something about classifying index
schemes are \textit{Region$_N$}?}
\item (\textbf{Index schemes to syntax}) \dnote{Mistral todo}
\end{enumerate}

### 2017-04-17 oopsla17/oopsla17.tex
moving checking/analysis into separate file
on inference part 1

  We also add the distinguished interval $[-\infty, \infty]$, which is simply an
  alias for \bz{}, but this notation prevents handling infinite interval
  separately in the following definitions, theorems, and proofs. Here, $-\infty$
  and $\infty$ behaves like top and bottom elements to \bz{} respectively.

@@

\begin{defn}[Subscript scheme]
  A \emph{subscript scheme} is an index scheme scheme where:
  %
  \begin{equation*}
    \forall i.\ 1 \leq i \leq N \implies
      \pi_i (S) = \{ p \}
      \; \vee \;
      \pi_i (S) = [-\infty, \infty]
    \end{equation*}
%
  That is, the $i^{th}$ component of the set is either a singleton in \bz{} or
  the infinite interval.
\end{defn}
%
\dnote{May want a predicate as well}

@@

\section{Analysis, Checking, and Inference}

\section{Evaluation}

>>>>

  We also add the distinguished interval $\interv{-\infty}{\infty}{}$, which is simply an
  alias for \bz{}, but this notation prevents handling infinite interval
  separately in the following definitions, theorems, and proofs. Here, $-\infty$
  and $\infty$ behaves like top and bottom elements to \bz{} respectively.

@@

\begin{defn}[Subscript scheme]
  A \emph{subscript scheme} is an index scheme scheme where:
  %
  \begin{equation*}
    \forall i.\ 1 \leq i \leq N \implies
      \pi_i (S) = \{ p \}
      \; \vee \;
      \pi_i (S) = \interv{-\infty}{\infty}{}
    \end{equation*}
%
  That is, the $i^{th}$ component of the set is either a singleton in \bz{} or
  the infinite interval.
\end{defn}
%
\dnote{May want a predicate as well}

@@

\section{Analysis, Checking, and Inference}
\label{sec:algorithms}
\input{algorithms}

\section{Evaluation}
\label{sec:evaluation}
